---
title: "Adam_Vuinovic_Final_Project_DDS"
output: html_document
date: "2023-02-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read in data and R libraries
```{r pressure, echo=FALSE}
# read in csv's and libraries

fullData = read.csv(file.choose(), sep = ",", header = T, quote = "")
dataNoAttrition = read.csv(file.choose(), sep = ",", header = T, quote = "")
dataNoSalary = read.csv(file.choose(), sep = ",", header = T, quote = "")


library(tidyr)
library(plyr)
library(jsonlite)
library(dplyr)
library(tidyverse)
library(class)
library(ggplot2)
library(caret)
library(e1071)
```

Exploratory data to understand composition of datasets
```{r pressure, echo=FALSE}

colSums(is.na(fullData)) #no null values

hist(fullData$Age)
hist(fullData$YearsAtCompany)
scatter.smooth(fullData$DistanceFromHome, 
               xlab = 'EmployeeID',
               ylab = 'Distance from Home [mi]')

#check unique values of standardHours
unique_values_stanHrs<- unique(fullData$StandardHours)

# print the result
print(unique_values_stanHrs)
#all the same value, we'll drop


#check unique values of overtime
unique_values_ot<- unique(fullData$OverTime)

# print the result
print(unique_values_ot)

```

Prep Data
```{r pressure, echo=FALSE}

#Filter out columns that won't be used to prevent overfitting model
preppedData = fullData %>% select(-YearsInCurrentRole, -Over18, -StandardHours, 
                                     -MonthlyRate, -DailyRate, -HourlyRate, -EducationField, -ID,
                                     -EmployeeNumber, -EmployeeCount, -TotalWorkingYears, -JobRole,
                                  -YearsWithCurrManager, -NumCompaniesWorked,-Education, -YearsAtCompany, 
                                  -TrainingTimesLastYear, -YearsSinceLastPromotion)

preppedData$MonthlyIncome = ifelse(preppedData$MonthlyIncome <= 5500, "Low", 
                                           ifelse(preppedData$MonthlyIncome <= 13000, "Medium", "High"))

```


Normalize ranges, One Hot Encode Categorical Variables, and convert back to data frame
```{r pressure, echo=FALSE}
# Convert to matrix to one hot encode categorical variables
design_matrix = model.matrix(preppedData$Attrition ~ ., data = preppedData)

# Normalize the data
#design_matrix_normalized <- scale(design_matrix)

# Convert the design matrix to a data frame
design_data <- data.frame(design_matrix, Attrition = preppedData$Attrition)


```


Create Test and Training Sets
```{r pressure, echo=FALSE}


# Split the data into training and test sets
set.seed(300)
train_index <- createDataPartition(design_data$Attrition, p = 0.6, list = FALSE)
training_data <- design_data[train_index, ]
test_data <- design_data[-train_index, ]

```


```{r pressure, echo=FALSE}
# Build the Naive Bayes model
model_nb <- naiveBayes(training_data[, -ncol(training_data)], training_data$Attrition)
```

```{r pressure, echo=FALSE}
# Predict the target variable for the test data Naive. Bayes
predictions_nb <- predict(model_nb, test_data[, -ncol(test_data)])

# Compare the predicted values with the actual values
table(predictions_nb, test_data$Attrition)
```

```{r pressure, echo=FALSE}
# Factor Levels of $Attrition
test_confusionmatrix = as.factor(test_data$Attrition)

# Full Confusion Matrix Evaluation
confusionMatrix(predictions_nb, test_confusionmatrix)

```




```{r pressure, echo=FALSE}
# Build the k-NN model
model_knn <- knn(train = training_data[, -ncol(training_data)],
                 test = test_data[, -ncol(test_data)],
                 cl = training_data$Attrition,
                 k = 5)
```


```{r pressure, echo=FALSE}
# convert model_knn to double
model_knn_dub = as.numeric(model_knn)

test_data_noAtt = test_data %>% select(-Attrition)
as.numeric(unlist(test_data_noAtt))

# Predict the target variable for the test data
predictions_knn <- predict(model_knn_dub, test_data_noAtt)

# Compare the predicted values with the actual values
table(predictions_knn, test_data$Attrition)

confusionMatrix_knn(predictions_knn, test_data$Attrition)
```


test_data$results = predictions_nb

ggplot(test_data, aes(x = BusinessTravelTravel_Frequently, fill = results)) + 
  geom_bar(position = "dodge") + 
  scale_fill_manual(values = c("green", "red")) + 
  ggtitle(" Age") + 
  xlab("Age") + 
  ylab("Count")
  
  